Estimating Information Transfer in Hard Parameter Sharing via Precise Bias-variance Asymptotics

### 2021-07-15 revision_journal/results.tex
minor fix
finish sec 5

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/verify_covariate_shift.eps}
		\caption{Verify variance asymptotics in finite dimensions}
		\label{fig_sec3_verify_cov}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_shift.eps}
		\caption{Sample size ratios vs. covariate shift}
		\label{fig_sec3_covariate}
	\end{subfigure}
	\caption{Figure \ref{fig_sec3_covariate}: we verify that Theorem \ref{thm_main_RMT} provides incredibly accurate estimates of the empirical variance under various finite sample sizes and dimensions.
	Figure \ref{fig_sec3_covariate}: We illustrate an intriguing dichotomy between covariate shift and sample size ratio: The dimension $p = 200$, $n_2 = 800$, and the noise variance $\sigma^2 = 1/4$.}
	\label{fig_sec31}
\end{figure}

@@

    We first describe a setting for modeling covariate shift.
    Let $\cS$ be the set of positive definite matrices $M^{\top} M$ such that for $i = 1,2,\dots,\floor{\frac p 2}$, $\lambda_{p - i}  = \frac 1 {\lambda_i}$, where $\lambda_1 \ge \lambda_2 \ge \cdots \ge \lambda_p$ are the eigenvalues of $M$.
    Under this setting, we ask: suppose that the sample sizes $n_1$ and $n_2$ are both fixed, which $M$ provides the best ``transfer'' in the sense of HPS estimator's excess risk?
    We observe a dichotomy that depends on whether or not $n_1$ is greater than $n_2$.
    \begin{claim}[Sample sizes vs. covariate shift]\label{claim_dichotomy}
        Let $g(M) = \frac{\sigma^2}{n_1 + n_2} \bigtr{\frac 1 {\alpha_1 M^{\top} M + \alpha_2}}$.
        Suppose that $n_1$ and $n_2$ are both fixed.
        Within the set of all possible covariate shift matrix $M\in\cS$, the following dichotomy holds:
        \begin{enumerate}
	        \item[i)] If $n_1 \ge n_2$, then $g(M)$ is minimized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
	        \item[ii)] If $n_1 < n_2$, then $g(M)$ is maximized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
        \end{itemize}
    \end{claim}

@@

Figure \ref{fig_sec3_covariate} illustrates a special case where $\lambda_1 = \cdots = \lambda_{\floor{p /2}} = \lambda > 0$ for $M^{\top}M$ and the rest of the eigenvalues are all equal to $1 / \lambda$.
Thus, $\lambda > 0$ captures the degree of covariate shift: higher $\lambda$ implies worse covariate shift.
We observe that our theoretical estimates using Corollary \ref{cor_hps_cov} matches the empirical risks incredibly accurately.
As a result, we indeed observe the dichotomy shown in Claim \ref{claim_dichotomy}.
Furthermore, for higher $\lambda$, HPS' excess risk decreases slower--indicating a worse rate of ``transfer'' from task one. 
As a remark, the classical work of \citet{david2010impossibility} have likewise shown impossibility results for transfer learning under covariate shift in a classification setting.
By contrast, our analysis applies to a (high-dimensional) regression setting.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_positive.eps}
		\caption{Positive transfer}
		\label{fig_sec3_model_positive}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_transition.eps}
		\caption{From positive to negative transfer}
		\label{fig_sec3_model_transition}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_negative.eps}
		\caption{Negative transfer}
		\label{fig_sec3_model_negative}
	\end{subfigure}	
	\caption{We illustrate three regimes of information transfer in HPS under model shift. When $\mu$ is small, HPS outperforms the OLS estimator irrespective of $n_1$. When $\mu$ is large, HPS performs worse than OLS. For an intermediate range of $\mu$, HPS outperforms OLS only for a restricted range of $n_1$. See Claim \ref{claim_model_shift} for the precise range of each regime. This simulation uses $p = 100, n_2 = 300, \sigma = 1/2$.}
	\label{fig_sec3_model_shift}
\end{figure}

@@

\todo{describe comparison to the TL estimator and the weighted OLS estimator.}

>>>>

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/verify_covariate_shift.eps}
		\caption{Illustrating variance estimates for finite $p$}
		\label{fig_sec3_verify_cov}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_shift.eps}
		\caption{Sample sizes vs. covariate shift}
		\label{fig_sec3_covariate}
	\end{subfigure}
	\caption{We verify that Theorem \ref{thm_main_RMT} provides incredibly accurate estimates of the empirical variance under various finite sample sizes and dimensions.
	Furthermore, we illustrate an intriguing dichotomy between sample sizes and covariate (cf. Claim \ref{claim_dichotomy}).
	When $n_1 \ge n_2$, the lowest risk for predicting task two's labels is achieved by transferring from a dataset that has the same covariance matrix as task two. 
	When $n_1 < n_2$, the lowest risk is instead achieved by transferring a covariate-shifted dataset.
	Both simulation uses $\sigma = 1/2$.
	Figure \ref{fig_sec3_verify_cov} fixes $\lambda = 4$ and varies $n_1, p$.
	Figure \ref{fig_sec3_covariate} fixes $p = 100, n_2 = 300$ while varying $n_1, \lambda$.}
	\label{fig_sec31}
\end{figure}

@@

    We first describe a setting for modeling covariate shift.
    Let $\cS$ be the set of positive definite matrices $M^{\top} M$ such that for $i = 1,2,\dots,\floor{\frac p 2}$, $\lambda_{p - i}  = \frac 1 {\lambda_i}$, where $\lambda_1 \ge \lambda_2 \ge \cdots \ge \lambda_p$ are the eigenvalues of $M$.
    Under this setting, we ask: suppose that the sample sizes $n_1$ and $n_2$ are both fixed, which $M$ provides the best ``transfer'' in the sense of HPS estimator's excess risk?
    We observe a dichotomy that depends on whether or not $n_1$ is greater than $n_2$.
    \begin{claim}[Sample sizes vs. covariate shift]\label{claim_dichotomy}
        Let $g(M) = \frac{\sigma^2}{n_1 + n_2} \bigtr{\frac 1 {\alpha_1 M^{\top} M + \alpha_2}}$.
        Suppose that $n_1$ and $n_2$ are both fixed.
        Within the set of all possible covariate shift matrix $M\in\cS$, the following dichotomy holds:
        \begin{enumerate}
	        \item[i)] If $n_1 \ge n_2$, then $g(M)$ is minimized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
	        \item[ii)] If $n_1 < n_2$, then $g(M)$ is maximized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
        \end{enumerate}
    \end{claim}

@@

Figure \ref{fig_sec3_covariate} illustrates a special case where $\lambda_1 = \cdots = \lambda_{\floor{p /2}} = \lambda > 0$ and the rest of the eigenvalues are all equal to $1 / \lambda$.
Thus, $\lambda > 0$ captures the degree of covariate shift: higher $\lambda$ implies worse covariate shift.
We observe that our theoretical estimates using Corollary \ref{cor_hps_cov} matches the empirical risks incredibly accurately.
As a result, we indeed observe the dichotomy shown in Claim \ref{claim_dichotomy}.
Furthermore, for higher $\lambda$, HPS' excess risk decreases slower--indicating a worse rate of ``transfer'' from task one. 
As a remark, the classical work of \citet{david2010impossibility} have likewise shown impossibility results for transfer learning under covariate shift in a classification setting.
By contrast, our analysis applies to a (high-dimensional) regression setting.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_positive.eps}
		\caption{Positive transfer}
		\label{fig_sec3_model_positive}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_transition.eps}
		\caption{From positive to negative transfer}
		\label{fig_sec3_model_transition}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/model_shift_negative.eps}
		\caption{Negative transfer}
		\label{fig_sec3_model_negative}
	\end{subfigure}	
	\caption{We illustrate three regimes of information transfer in HPS under model shift. When $\mu$ is small, HPS outperforms the OLS estimator irrespective of $n_1$. When $\mu$ is large, HPS performs worse than OLS. For an intermediate range of $\mu$, HPS outperforms OLS only for a restricted range of $n_1$. See Claim \ref{claim_model_shift} for the precise range of each regime. This simulation fixes $p = 100, n_2 = 300, \sigma = 1/2$ while varying $n_1, \mu$.}
	\label{fig_sec3_model_shift}
\end{figure}

@@
